---
title: "Identification of key risk factors associated with severe hypoglycemia in older
  adults with Type 1 Diabetes"
author: "Nikki L. B. Freeman"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

This notebook contains the main steps in the analysis for "Identification of key risk factors associated with severe hypoglycemia in older adults with Type 1 Diabetes". Function files are located in the same directory as this notebook. 

# Packages, functions, and data

The `getRawDataTables` function reads in the individual data tables from the raw data set. 

```{r, message = FALSE, warning = FALSE, error = FALSE}
# Packages needed for the analysis
library(tidyverse)
library(mice)
library(caret)
library(randomForest)
library(glmnet)

# Functions
source("1_createAnalysisRaw.R")
source("2_dataPreparation2.R")
source("3_imputation.R")
source("4_randomForest.R")
source("5_regLogisticReg.R")

# Load the raw data
dataTables <- getRawDataTables("../")
```

# Data cleaning and preparation

The `collectVariables` function extracts from each data table, the variables of interest for the analysis. For some variables, minor data cleaning is completed, including 

  * Adding an inclusion/exclusion variable: two participants are excluded due to not having any demographic data available
  * Adjudicating duplicate lab results: if one entry is NULL and the other is not, take the non-NULL entry; otherwise, average the entries.
  * Scoring the diabetes numeracy test
  * Scoring the geriatric depression test
  * Identifying beta blockers from the medication list and creating a binary beta blocker variable
  * Scoring the hypoglycemia unawareness survey
  
The`makeAnalysisDataWithMissing`

```{r}
# Collect the variables of interest from the various tables and put into a single analysisData_raw dataframe
# Some variables are missing their scores or are re-coded
analysisData_raw <- collectVariables(dataTables)
# Write the analysis data to a csv
write_csv(x = analysisData_raw, file = "../2_pipeline/1_analysis_raw.csv")

# Create analytic data set for analyis
analysisWithMissing <- makeAnalysisDataWithMissing(analysisData_raw = analysisData_raw)
# Write the analysis data set to a csv
write_csv(x = analysisWithMissing, file = "../2_pipeline/2_analysisDataWithMissing.csv")
```

# Imputation and preparation of testing and training sets
```{r, message = FALSE, warning = FALSE, results = 'hide'}
set.seed(5678)
imputedData <- getImputationDataSets(analysisWithMissing = analysisWithMissing)
trainTest <- createTrainTest(analysisWithMissing = analysisWithMissing, imputedData = imputedData)

```

# Random forest
## Model fitting
```{r}

rfObjs <- map(trainTest$trainingSets_list, rfHelper)

rfPredictions <- map(rfObjs, predict, newdata = trainTest$testingData %>% 
                       dplyr::select(-c(BCaseControlStatus, PtID)))

rfImportance <- map(rfObjs, rfImportanceHelper)

```

## Model assessment

```{r}
rf_metrics <- map(rfPredictions, confusionMatrix, reference = factor(trainTest$testingData$BCaseControlStatus))

rf_plotMetrics <- data.frame(reduce(map(1:5, ~rf_metrics[[.]]$byClass), rbind)) %>%
  mutate(Method = "RF")
```


# L1-Regularized logistic regression
## Model fitting
```{r}

regLogisticRegObj <- map(trainTest$trainingSets_list, regLogisticReg_helper)  

regLogisticRegPredictions <- map(regLogisticRegObj, predict, convertTestCategoricalToBinary(trainTest$testingData))

regLogisticRegImportance <- map(regLogisticRegObj, varImp)

```


## Model assessment
```{r}
regLog_metrics <- map(regLogisticRegPredictions, confusionMatrix, reference = factor(trainTest$testingData$BCaseControlStatus))
regLog_plotMetrics <- data.frame(reduce(map(1:5, ~regLog_metrics[[.]]$byClass), rbind)) %>%
  mutate(Method = "LR")
```


# SVM

## Model fitting
```{r}

svmHelper <- function(trainingSet){
  temp <- convertTrainingCategoricalToBinary(trainingSet)
  e1071::svm(x = temp, y = as.factor(trainingSet$BCaseControlStatus))
}

svmObjs <- map(trainTest$trainingSets_list, svmHelper)

svmPredictions <- map(svmObjs, predict, convertTestCategoricalToBinary(trainTest$testingData))

# svmImportance <- map(svmObjs, varImp, scale = FALSE)
```


## Model assessment
```{r}
svm_metrics <- map(svmPredictions, confusionMatrix, reference = factor(trainTest$testingData$BCaseControlStatus))
svm_plotMetrics <- data.frame(reduce(map(1:5, ~svm_metrics[[.]]$byClass), rbind)) %>%
  mutate(Method = "SVM")
```


# Prediction performance plots
```{r}
predPerfDF <- bind_rows(rf_plotMetrics, regLog_plotMetrics, svm_plotMetrics) %>%
  mutate(fold = rep(1:5, 3)) %>%
  dplyr::select(Method, fold, Sensitivity, Specificity, Precision, Recall) %>% 
  pivot_longer(cols = c(Sensitivity, Specificity, Precision, Recall), 
               names_to = "metric", values_to = "value") %>%
  replace_na(list(value = 0))

predPerfPlot <- predPerfDF %>% 
  ggplot(aes(x = Method, y = value, color = Method)) +
  geom_point() +
  facet_wrap(~metric) +
  theme_minimal() +
  labs(y = element_blank(), x = element_blank()) +
  theme(legend.position = "bottom")

ggsave(predPerfPlot, file = "../3_output/0_predPerfPlot.png", device = "png",
       bg = "white", width = 12, height = 7, units = "in")
```

# Average predictive performance
```{r}
# Average of the predictive performance measures across folds
predPerfDF %>%
  group_by(Method, metric) %>%
  summarise(mean = round(mean(value), 2),
            sd = round(sd(value), 2)) %>%
  arrange(metric)

# Average accuracy, etc across folds
round(colMeans(reduce(map(rf_metrics, ~pluck(., "overall")), rbind)), 2)

round(colMeans(reduce(map(regLog_metrics, ~pluck(., "overall")), rbind)), 2)

round(colMeans(reduce(map(svm_metrics, ~pluck(., "overall")), rbind)), 2)
```


# Variable importance

```{r}
data.frame(reduce(map2(1:5,rfImportance, function(x, y) y %>% mutate(fold = x)), rbind)) %>%
  arrange(fold, desc(MeanDecreaseGini)) %>%
  group_by(fold) %>%
  slice_max(MeanDecreaseGini, n = 5)

regLogisticRegImportance
```


