---
title: "Identification of key risk factors associated with severe hypoglycemia in older
  adults with Type 1 Diabetes"
author: "Nikki L. B. Freeman"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

This notebook contains the main steps in the analysis for "Identification of key risk factors associated with severe hypoglycemia in older adults with Type 1 Diabetes". Function files are located in the same directory as this notebook. 

# Packages, functions, and data

The `getRawDataTables` function reads in the individual data tables from the raw data set. 

```{r, message = FALSE, warning = FALSE, error = FALSE}
# Packages needed for the analysis
library(tidyverse)
library(mice)
library(caret)
library(randomForest)
library(glmnet)
library(xtable)
library(table1)

# Functions
source("1_createAnalysisRaw.R")
source("2_dataPreparation2.R")
source("3_imputation.R")
source("4_randomForest.R")
source("5_regLogisticReg.R")

# Load the raw data
dataTables <- getRawDataTables("../")
```

# Data cleaning and preparation

The `collectVariables` function extracts from each data table, the variables of interest for the analysis. For some variables, minor data cleaning is completed, including 

  * Adding an inclusion/exclusion variable: two participants are excluded due to not having any demographic data available
  * Adjudicating duplicate lab results: if one entry is NULL and the other is not, take the non-NULL entry; otherwise, average the entries.
  * Scoring the diabetes numeracy test
  * Scoring the geriatric depression test
  * Identifying beta blockers from the medication list and creating a binary beta blocker variable
  * Scoring the hypoglycemia unawareness survey
  
The`makeAnalysisDataWithMissing` does additional data cleaning and perparation to yield a dataframe with all of the observed data that will be used for the analysis. The cleaning steps include: 
  
  * Re-coding education into categories (highest educational attainment <= HS, Some college or college degree, or advanced degree)
  * Re-coding insurance into a binary variable for having insurance (public, private, single service) or not
  * Calculating BMI and categorizing into BMI classes (BMI < 18.5 as underweight, 18.5 $\le$ BMI < 25 as normal weight, 15 $\le$ BMI < 30 as overweight, and BMI $\ge$ 30 as obese)
  * Re-coding total insulin units as <40, $\ge$ 40 and <60, and $\ge$ 60
  * Averaging the two frailty walk scores to  yield a single frailty walk score for each individual

```{r}
# Collect the variables of interest from the various tables and put into a single analysisData_raw dataframe
# Some variables are missing their scores or are re-coded
analysisData_raw <- collectVariables(dataTables)
# Write the analysis data to a csv
write_csv(x = analysisData_raw, file = "../2_pipeline/1_analysis_raw.csv")

# Create analytic data set for analyis
analysisWithMissing <- makeAnalysisDataWithMissing(analysisData_raw = analysisData_raw)
# Write the analysis data set to a csv
write_csv(x = analysisWithMissing, file = "../2_pipeline/2_analysisDataWithMissing.csv")
```

# Imputation and preparation of testing and training sets

The `getImputationDataSets` function takes the observed data for analysis and creates 5 imputed dated sets. The full observed data set for imputation. The `createTrainTest` function takes the imputed dated and observed data to generate a test set composed of only complete cases and 5 imputation sets that are used for training the models (effectively, these are folds).

```{r, message = FALSE, warning = FALSE, results = 'hide'}
set.seed(5678)
imputedData <- getImputationDataSets(analysisWithMissing = analysisWithMissing)
trainTest <- createTrainTest(analysisWithMissing = analysisWithMissing, imputedData = imputedData)

```

# Random forest
## Model fitting

This code chunk fits a RF to the training sets, generates predictions for the testing set, and extracts the variable importance for the trained models.

```{r}

rfObjs <- map(trainTest$trainingSets_list, rfHelper)

rfPredictions <- map(rfObjs, predict, newdata = trainTest$testingData %>% 
                       dplyr::select(-c(BCaseControlStatus, PtID)))

rfImportance <- map(rfObjs, rfImportanceHelper)

```

## Model assessment

This chunk grabs the prediction metrics for the RF model and creates a dataframe of the prediction metrics that will be used for the prediction performance plot.

```{r}
rf_metrics <- map(rfPredictions, confusionMatrix, reference = factor(trainTest$testingData$BCaseControlStatus))

rf_plotMetrics <- data.frame(reduce(map(1:5, ~rf_metrics[[.]]$byClass), rbind)) %>%
  mutate(Method = "RF")
```


# L1-Regularized logistic regression
## Model fitting

This code chunk fits a LR model to the training sets, generates predictions for the testing set, and extracts the variable importance for the trained models.

```{r}

regLogisticRegObj <- map(trainTest$trainingSets_list, regLogisticReg_helper)  

regLogisticRegPredictions <- map(regLogisticRegObj, predict, convertTestCategoricalToBinary(trainTest$testingData))

regLogisticRegImportance <- map(regLogisticRegObj, varImp)

```


## Model assessment

This chunk grabs the prediction metrics for the LR model and creates a dataframe of the prediction metrics that will be used for the prediction performance plot.

```{r}
regLog_metrics <- map(regLogisticRegPredictions, confusionMatrix, reference = factor(trainTest$testingData$BCaseControlStatus))
regLog_plotMetrics <- data.frame(reduce(map(1:5, ~regLog_metrics[[.]]$byClass), rbind)) %>%
  mutate(Method = "LR")
```


# SVM

## Model fitting

This code chunk fits a SVM model to the training sets and generates predictions for the testing set.

```{r}

svmHelper <- function(trainingSet){
  temp <- convertTrainingCategoricalToBinary(trainingSet) %>%
    select(-race_asian)
  e1071::svm(x = temp, y = as.factor(trainingSet$BCaseControlStatus))
}

svmObjs <- map(trainTest$trainingSets_list, svmHelper)

svmPredictions <- map(svmObjs, predict, convertTestCategoricalToBinary_svm(trainTest$testingData))

```


## Model assessment

This chunk grabs the prediction metrics for the LR model and creates a dataframe of the prediction metrics that will be used for the prediction performance plot.

```{r}
svm_metrics <- map(svmPredictions, confusionMatrix, reference = factor(trainTest$testingData$BCaseControlStatus))
svm_plotMetrics <- data.frame(reduce(map(1:5, ~svm_metrics[[.]]$byClass), rbind)) %>%
  mutate(Method = "SVM")
```


# Prediction performance plots

This chunk generates the predictive performance plots. 

```{r}
predPerfDF <- bind_rows(rf_plotMetrics, regLog_plotMetrics, svm_plotMetrics) %>%
  mutate(fold = rep(1:5, 3)) %>%
  dplyr::select(Method, fold, Sensitivity, Specificity, Precision, Recall) %>% 
  pivot_longer(cols = c(Sensitivity, Specificity, Precision, Recall), 
               names_to = "metric", values_to = "value") %>%
  replace_na(list(value = 0))

predPerfPlot <- predPerfDF %>% 
  ggplot(aes(x = Method, y = value, color = Method)) +
  geom_point() +
  facet_wrap(~metric) +
  theme_minimal() +
  labs(y = element_blank(), x = element_blank()) +
  theme(legend.position = "bottom")

ggsave(predPerfPlot, file = "../3_output/0_predPerfPlot.png", device = "png",
       bg = "white", width = 12, height = 7, units = "in")
```

# Average predictive performance
```{r}
# Average of the predictive performance measures across folds
predPerfDF %>%
  group_by(Method, metric) %>%
  summarise(mean = round(mean(value), 2),
            sd = round(sd(value), 2)) %>%
  arrange(metric)

# Average accuracy, etc across folds
round(colMeans(reduce(map(rf_metrics, ~pluck(., "overall")), rbind)), 2)

round(colMeans(reduce(map(regLog_metrics, ~pluck(., "overall")), rbind)), 2)

round(colMeans(reduce(map(svm_metrics, ~pluck(., "overall")), rbind)), 2)
```


# Variable importance

```{r}
data.frame(reduce(map2(1:5,rfImportance, function(x, y) y %>% mutate(fold = x)), rbind)) %>%
  arrange(fold, desc(MeanDecreaseGini)) %>%
  group_by(fold) %>%
  slice_max(MeanDecreaseGini, n = 5)

regLogisticRegImportance
```

# Descriptives table 

```{r}
table1df <- analysisWithMissing %>%
  filter(str_detect(exclude, "No")) %>%
  mutate(Gender = factor(Gender, levels = c("F", "M"), labels = c("Female", "Male")),
         Race = factor(Race, 
                       levels = c("White", "Asian", "Unknown/not reported", 
                                  "Black/African American", 
                                  "American Indian/Alaskan Native"),
                       labels = c("White", "Asian", "Unknown/not reported", 
                                  "Black/African American", 
                                  "American Indian/Alaskan Native")),
         Ethnicity = factor(Ethnicity),
         `Exercise (days per week)` = DaysWkEx, 
         `Lives alone` = LiveAlone,
         `Insulin delivery method` = InsDeliveryMethod,
         `Units Insulin (bolus or long acting)` = UnitsInsBasalOrLongAct,
         `Number boluses or injections short acting per day` = NumPumpBolusOrShortAct,
         `Days hospitalized DKA` = NumHospDKA,
         HBA1C = as.numeric(HBA1C),
         `Beta blocker use` = factor(betaBlocker, levels = c(1, 0), labels = c("Yes", "no")),
         `Symbolic Digits Written test` = SymbDigWTotCorr,
         `Symbolic Digits Oral test` = SymbDigOTotCorr,
         `Trail Making test` = TrailMakATotTime,
         `Grooved Peg Board test (dominant hand)` = GrPegDomTotTime,
         `Functional Activity score` = FuncActTotTestScore,
         `Duke Social Support scale` = DukeSocTotDSSI,
         `Geriatric Depression score` = geriDepressionScore,
         `Hypoglycemia unawareness` = factor(hypoUnaware, levels = c("hypo aware", "reduced awareness", "unaware"),
                                             labels = c("Aware", "Reduced awareness", "Unaware")),
         `Diabetes Numeracy score` = diabNumeracyScore,
         `Montreal Cognitive Assessment score` = MoCATotal,
         `Education` = factor(educationCat, levels = c("HS", "College", "Advanced degree"),
                              labels = c("<= HS", "Any college", "Advanced degree")),
         `Any insurance` = factor(anyInsurance, levels = c(0, 1),
                                  labels = c("No", "Yes")),
         `Alcohol use (days per month)` = DaysMoDrinkAlc,
         `Insulin dose` = factor(insDoseCat, levels = c("less40", "40to60", "greater60"),
                                 labels = c("<40", "40-60", ">60")),
         `Detectable C-peptide` = factor(detectableCPEP, levels = c(0, 1),
                                         labels = c("<0.017", ">=0.017")),
         `Abnormal creatinine` = factor(abnormalCreatinine, levels = c(0, 1),
                                        labels = c("<=1.1 females/<=1.2 males", ">1.1 females/>1.2 males")),
         `Average frailty walk time` = frailty,
         `BMI category` = factor(bmiCat, levels = c("underweight", "normal", "overweight", "obese"),
                                 labels = c("Underweight", "Normal weight", "Overweight", "Obese"))) 

table1_out <- table1(~ Gender + Race + Ethnicity + 
         `Education` + `Any insurance` + `BMI category` +
         `Exercise (days per week)` +
         `Lives alone` + `Alcohol use (days per month)` +
         `Insulin delivery method` + `Insulin dose` +
         `Number boluses or injections short acting per day` +
         `Days hospitalized DKA` + HBA1C + `Detectable C-peptide` + 
         `Abnormal creatinine` + `Beta blocker use` +
         `Symbolic Digits Written test` + `Symbolic Digits Oral test` +
         `Trail Making test` + `Grooved Peg Board test (dominant hand)` +
         `Functional Activity score` + `Duke Social Support scale` +
         `Geriatric Depression score` + `Hypoglycemia unawareness` +
         `Diabetes Numeracy score` + `Montreal Cognitive Assessment score` +
         `Average frailty walk time`
         | BCaseControlStatus, data = table1df)
# Knitted table
table1_out

# Create the latex table
print(xtable(as.data.frame(table1_out)), include.rownames = FALSE)
```

