---
title: "Identification of key risk factors associated with severe hypoglycemia in older
  adults with Type 1 Diabetes"
author: "Nikki L. B. Freeman"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

This notebook contains the main steps in the analysis for "Identification of key risk factors associated with severe hypoglycemia in older adults with Type 1 Diabetes". Function files are located in the same directory as this notebook. 

# Packages, functions, and data
```{r, message = FALSE, warning = FALSE, error = FALSE}
# Packages needed for the analysis
library(tidyverse)
library(mice)
library(caret)
library(randomForest)
library(glmnet)

# Functions
source("1_createAnalysisRaw.R")
source("2_dataPreparation2.R")
source("3_imputation.R")
source("4_randomForest.R")
source("5_regLogisticReg.R")

# Load the raw data
dataTables <- getRawDataTables("../")
```

# Data cleaning and preparation
```{r}
# Collect the variables of interest from the various tables and put into a single analysisData_raw dataframe
# Some variables are missing their scores or are re-coded
analysisData_raw <- collectVariables(dataTables)
# Write the analysis data to a csv
write_csv(x = analysisData_raw, file = "../2_pipeline/1_analysis_raw.csv")

# Create analytic data set for analyis
analysisWithMissing <- makeAnalysisDataWithMissing(analysisData_raw = analysisData_raw)
# Write the analysis data set to a csv
write_csv(x = analysisWithMissing, file = "../2_pipeline/2_analysisDataWithMissing.csv")
```

# Imputation and preparation of testing and training sets
```{r, message = FALSE, warning = FALSE, results = 'hide'}
set.seed(5678)
imputedData <- getImputationDataSets(analysisWithMissing = analysisWithMissing)
trainTest <- createTrainTest(analysisWithMissing = analysisWithMissing, imputedData = imputedData)

```

# Random forest
## Model fitting
```{r}

rfObjs <- map(trainTest$trainingSets_list, rfHelper)

rfPredictions <- map(rfObjs, predict, newdata = trainTest$testingData %>% 
                       dplyr::select(-c(BCaseControlStatus, PtID)))

rfImportance <- map(rfObjs, rfImportanceHelper)

```

## Model assessment

```{r}
map(rfPredictions, confusionMatrix, reference = factor(trainTest$testingData$BCaseControlStatus))
```


# L1-Regularized logistic regression
## Model fitting
```{r}

regLogisticRegObj <- map(trainTest$trainingSets_list, regLogisticReg_helper)  

regLogisticRegPredictions <- map(regLogisticRegObj, predict, convertTestCategoricalToBinary(trainTest$testingData))

regLogisticRegImportance <- map(regLogisticRegObj, varImp)

```


## Model assessment
```{r}
map(regLogisticRegPredictions, confusionMatrix, reference = factor(trainTest$testingData$BCaseControlStatus))
```


# SVM

## Model fitting
```{r}

svmHelper <- function(trainingSet){
  temp <- convertTrainingCategoricalToBinary(trainingSet)
  e1071::svm(x = temp, y = as.factor(trainingSet$BCaseControlStatus))
}

svmObjs <- map(trainTest$trainingSets_list, svmHelper)

svmPredictions <- map(svmObjs, predict, convertTestCategoricalToBinary(trainTest$testingData))

# svmImportance <- map(svmObjs, varImp, scale = FALSE)
```


## Model assessment
```{r}
map(svmPredictions, confusionMatrix, reference = factor(trainTest$testingData$BCaseControlStatus))
```


