---
title: "SH Data Preparation"
output:
  html_document:
    df_print: paged
---

```{r chunk set up, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r "packages, scripts, and data", include = FALSE}
# Packages
library(tidyverse)

# Scripts
source("2_dataPreparation.R")

# Data
analysisRaw <- read_csv("../2_pipeline/1_analysis_raw.csv")
```

# Assess missingness
Summaries exclude the two study participants (PtIDs 194 and 183) that do not have any demographic data. They will not be included in the analysis.

## Demographics, lifestyle, and diabetes management

### Sex

```{r}
getOverallStratifiedCounts(analysisRaw, "Gender")
```
Sex is a complete variable; no additional cleaning.

### Race
```{r}
getOverallStratifiedCounts(analysisRaw, "Race")
```

```{r, include = FALSE}
analysisRaw %>% filter(str_detect(Race, 'Unknown'))
```
PtID 157: Race is "Unknown/not reported" and ethnicity is "Hispanic or Latino"


### Ethnicity
```{r}
getOverallStratifiedCounts(analysisRaw, "Ethnicity")
```
```{r, include = FALSE}
analysisRaw %>% filter(str_detect(Ethnicity, "Unknown"))
```
PtID 69: Race is white and ethnicity is Unknown/not reported

### Age
Still looking for this in the CRFs

### Disease duration
Propose defn: Age - Age dx; have age dx, still looking for age

### Age of T1 dx
```{r}
getOverallStratifiedCounts(analysisRaw, "T1DDiagAge")
```
None are missing

```{r, include= FALSE}
analysisRaw <- analysisRaw %>% select(-T1DDiagAgeUnk)

```



### Education
```{r}
getOverallStratifiedCounts(analysisRaw, "EduLevel")
getOverallStratifiedCounts(analysisRaw, "EduLevelNoAns")
getOverallStratifiedCounts(analysisRaw, "EduLevelUnk")
```

```{r, include = FALSE}
analysisRaw %>% filter(EduLevelNoAns == 1)
```
PtIDs 165 and 167: No answer to the education level question

Use the same education categories as Weinstock (2016): Less than HS or high school diploma/GED, some college/associate or bachelor degree, master/professional or doctorate degree

```{r}
analysisRaw %>% select(EduLevel) %>%
  distinct()

analysisRaw <- analysisRaw %>%
  mutate(educationCat = if_else(EduLevel %in% c("High school graduate/diploma/GED",
                                                "11th Grade",
                                                "12th Grade - no diploma",
                                                "7th or 8th Grade",
                                                "9th Grade"), "HS", "?")) %>%
  mutate(educationCat = if_else(EduLevel %in% c("Associate Degree",
                                                "Bachelor's Degree",
                                                "Some college but no degree"), 
                                "College", educationCat)) %>%
  mutate(educationCat = if_else(EduLevel %in% c("Doctorate Degree",
                                                "Master's Degree",
                                                "Professional Degree"),
                                "Advanced degree", educationCat)) %>%
  mutate(educationCat = if_else(EduLevelNoAns == 1 & !is.na(EduLevelNoAns), as.character(NA), educationCat)) %>%
  select(-c(EduLevel, EduLevelNoAns, EduLevelUnk))

getOverallStratifiedCounts(analysisRaw %>% filter(str_detect(exclude, "No")), "educationCat")
```
With the new education categories, PtIDs 165 and 167 continue to be missing

### Health Insurance
```{r}
analysisRaw %>% 
  filter(str_detect(exclude, "No")) %>%
  group_by(InsPriv, InsGov, InsSingleService, InsNoCoverage, InsUnkown, InsNoAns) %>%
  count()
```

```{r, include = FALSE}
analysisRaw %>% 
  rowwise() %>%
  mutate(anyInsurance = sum(c(InsPriv, InsGov, InsSingleService), na.rm = TRUE)) %>%
  select(-c(InsPriv, InsGov, InsSingleService)) %>%
  select(PtID, anyInsurance, InsNoCoverage, InsUnkown, InsNoAns) %>%
  filter(anyInsurance == 0)
```
PtIDs 19, 33, 1: No insurance coverage
PtID 74: Did not answer

Create a new variable for has insurance/does not have insurance coverage

```{r}
analysisRaw <- analysisRaw %>% 
  rowwise() %>%
  mutate(nInsurance = sum(c(InsPriv, InsGov, InsSingleService), na.rm = TRUE)) %>% 
  mutate(anyInsurance = if_else(nInsurance > 0, 1, -1),
         anyInsurance = if_else(!is.na(InsNoCoverage) & InsNoCoverage == 1, 0, anyInsurance),
         anyInsurance = if_else(InsNoAns == 1 & !is.na(InsNoAns), as.numeric(NA), anyInsurance)) %>%
  select(-c(InsPriv, InsGov, InsSingleService, InsNoAns, InsNoCoverage))  %>%
  ungroup()

getOverallStratifiedCounts(analysisRaw, "anyInsurance")
```
With the new variable `anyInsurance`, only PtID 74 is missing.

### BMI

Construct the BMI variable from height and weight (kg/m^2). Round to 1 decimal place as is customary in reporting.

```{r}
# construct the BMI variable
analysisRaw <- analysisRaw %>%
  # Convert pounds to kilograms
  mutate(bmiNumerator = if_else(WeightUnits == "kg", Weight, -1),
         bmiNumerator = if_else(WeightUnits == "lbs", Weight*0.453592, bmiNumerator),
         bmiDenominator = if_else(HeightUnits == "cm", (Height*0.01)^2, -1),
         bmiDenominator = if_else(HeightUnits == "in", (Height*0.0254)^2, bmiDenominator),
         bmi = round(bmiNumerator/bmiDenominator, 1)) %>%
  select(-c(Weight, WeightUnits, WeightUnk, Height, HeightUnk, HeightUnits, bmiNumerator, bmiDenominator))

analysisRaw %>%
  ggplot(aes(x = bmi)) +
  geom_histogram(bins = 10) +
  theme_minimal() +
  xlab("BMI") +
  ylab("Count")
```

PtIDs 124, 16, 30, and 112: Incomplete data for BMI calculation (All are missing height; 112 also missing weight)

As in Weinstock (2016), we'll also create a categorical bmi variable:  underweight (BMI <18.5), normal weight (18.5 <= BMI < 25), overweight (25 <= BMI < 30), obese (BMI >= 30)

```{r}
analysisRaw <- analysisRaw %>%
  mutate(bmiCat = if_else(bmi <18.5, "underweight", "?"),
         bmiCat = if_else(18.5 <= bmi & bmi < 25, "normal", bmiCat),
         bmiCat = if_else(25 <= bmi & bmi < 30, "overweight", bmiCat),
         bmiCat = if_else(bmi >= 30, "obese", bmiCat)) 

analysisRaw %>%
  filter(str_detect(exclude, "No")) %>%
  ggplot(aes(x = bmiCat)) +
  geom_bar() +
  scale_x_discrete(name = "BMI category",
                   breaks = c("underweight", "normal", "overweight", "obese", NA),
                   labels = c("Underweight", "Normal", "Overweight", "Obese", "Missing")) +
  theme_minimal() +
  ylab("Count")
```


### Exercise (days/week)
```{r}
analysisRaw %>% select(PtID, DaysWkEx, DaysWkExUnk, DaysWkExDNA) %>%
  ggplot(aes(x = DaysWkEx)) +
  geom_histogram(bins = 8) +
  xlab("Number of days") +
  ylab("Count")

```

```{r, include = FALSE}
analysisRaw %>% filter(DaysWkExUnk == 1) 
analysisRaw %>% filter(is.na(DaysWkEx))
analysisRaw <- analysisRaw %>% select(-c(DaysWkExDNA, DaysWkExUnk))
```
PtIDs 124 and 27: Exercise days per week is unknown

### Alcohol use

```{r}
analysisRaw <- analysisRaw %>%
  mutate(DaysMoDrinkAlc = if_else(!is.na(DaysWkDrinkAlcNone) & DaysWkDrinkAlcNone == 1, 0, DaysWkDrinkAlc)) %>%
  select(-c(DaysWkDrinkAlc, DaysWkDrinkAlcNone, DaysWkDrinkAlcUnk))

getOverallStratifiedCounts(analysisRaw, "DaysMoDrinkAlc")
```
```{r, include = FALSE}
analysisRaw %>% filter(is.na(DaysMoDrinkAlc))
```

PtID 195: is missing Days month drink alcohol

### Lives alone

```{r}
getOverallStratifiedCounts(analysisRaw, "LiveAlone")
```
This variable is complete

### Insulin pump use (Y/N)

```{r}
getOverallStratifiedCounts(analysisRaw, "InsDeliveryMethod")
```
This variable is complete.

### Daily dose

Following Weinstock (2016) we'll categorize units of total insulin as <40, 40-60, and >= 60
```{r}
analysisRaw <- analysisRaw %>%
  mutate(insDoseCat = if_else(UnitsInsTotal < 40, "less40", "?"),
         insDoseCat = if_else(UnitsInsTotal >= 40 & UnitsInsTotal < 60, "40to60", insDoseCat),
         insDoseCat = if_else(UnitsInsTotal >= 60, "greater60", insDoseCat)) %>%
  select(-c(UnitsInsTotalUnk, UnitsInsTotal))

getOverallStratifiedCounts(analysisRaw, "insDoseCat")
```
```{r, include = FALSE}
analysisRaw %>% filter(is.na(insDoseCat))
```
PtIDs 19, 196, 164, 84, 110, 91, 24, 149, 138, 113, 124, 27, 129, 109, 1, 98, 74, 126 are missing insulin dose

```{r}
# Analysis of basal/long acting and bolus/short acting
getOverallStratifiedCounts(analysisRaw, "UnitsInsBasalOrLongAct")
getOverallStratifiedCounts(analysisRaw, "NumPumpBolusOrShortAct")
```
```{r, include = FALSE}
analysisRaw %>% filter(is.na(UnitsInsBasalOrLongAct)) %>% select(PtID, UnitsInsBasalOrLongAct, UnitsInsBasalOrLongActUnk)
analysisRaw %>% filter(is.na(NumPumpBolusOrShortAct)) %>% select(PtID, NumPumpBolusOrShortAct, NumPumpBolusOrShortActUnk)
analysisRaw <- analysisRaw %>% select(-c(UnitsInsBasalOrLongActUnk, NumPumpBolusOrShortActUnk))
```
`UnitsInsBasalOrLongAct` is missing for PtIDs: 56, 181, 144, 91, 153, 131, 150, 140, 149, 14, 138, 113, 124, 191, 129, 12, 120, 35, 60, 18, 66, 15, 176, 193, 46, 147, 160

`NumPumpBolusOrShortAct` is missing for PtIDs: 181, 136, 81, 148, 184, 165, 110, 116, 91, 153, 150, 149, 14, 113, 124, 27, 191, 129, 12, 120, 60, 93, 156, 78, 185, 18, 66, 122, 74, 183, 193, 126, 160, 85

### Frequency of testing glucose

```{r}
getOverallStratifiedCounts(analysisRaw, "NumMeterCheckDay")
```
This variable is complete.

### HbA1c
```{r}
analysisRaw %>% 
  ggplot(aes(x = HBA1C)) +
  geom_histogram()
```

```{r, include = FALSE}
analysisRaw %>% select(PtID, HBA1C) %>% filter(is.na(HBA1C))

```
This variable is complete


### Detectable C-peptide
Table 2 says detectable C-peptide is >= 0.017
```{r}
# Make the detectable c-pep variable
analysisRaw <- analysisRaw %>% 
  mutate(detectableCPEP = if_else(CPEP != "<0.017", 1, 0)) %>% 
  select(-CPEP)

analysisRaw %>% ggplot(aes(x = detectableCPEP)) +
  geom_bar()
  
```

```{r, include = FALSE}
analysisRaw %>% filter(is.na(detectableCPEP)) %>% select(PtID)

```
Missing for PtIDs: 188, 74

### Beta blocker use

```{r}
analysisRaw %>%
  ggplot(aes(x = betaBlocker)) +
  geom_bar()
```
```{r}
analysisRaw %>% filter(is.na(betaBlocker))
```
No meds were recorded for 153 and 52; study completion form says no medications to be entered

### Abnormal creatinine
Table 2 says Abnormal creatinine defined as >1.1 mg/dL for females and >1.2 mg/dL for males

Waiting on definition
```{r}
# construct abnormalCreatine
analysisRaw <- analysisRaw %>% 
  mutate(abnormalCreatinine = if_else((`CREA-S` > 1.1 & Gender == "F") | (`CREA-S` > 1.2 & Gender == "M"), 1, 0)) %>%
  select(-`CREA-S`)
getOverallStratifiedCounts(analysisRaw, "abnormalCreatinine")
```

```{r, include = FALSE}
analysisRaw %>% filter(is.na(abnormalCreatinine)) %>% select(PtID)

```
PtIDs 188 and 74 are missing creatinine measures

### History of DKA

```{r}
getOverallStratifiedCounts(analysisRaw, "NumHospDKA")
```

```{r, include = FALSE}
analysisRaw %>% filter(is.na(NumHospDKA)) %>% select(PtID, NumHospDKA, NumHospDKAUnk)

analysisRaw <- analysisRaw %>% select(-NumHospDKAUnk)

getOverallStratifiedCounts(analysisRaw, "NumHospDKA")
```
PtID 143: Number of hosptal DKA is unknown

There's not a lot of variation here. Recode as 0 vs 1+? Still not going to have a lot of variation (e.g. only 2 among the controls)

## Cogntive assessments
### Montreal Cognitive Assessment

```{r}
getOverallStratifiedCounts(analysisRaw, "MoCATotal")
```

```{r, include = FALSE}
analysisRaw %>% filter(is.na(MoCATotal)) %>% select(PtID, MoCATotal, MoCANotDone)

analysisRaw <- analysisRaw %>% select(-MoCANotDone)
```
PtID 144: MoCA was not done

### Hopkins Verbal Learning Test - Total Recall

Need to learn how this is scored

### Hopkins Verbal Learning Test - Delayed Recall

Need to learn how this is scored

### Symbol Digit Modalities Test - Written

```{r}
getOverallStratifiedCounts(analysisRaw, "SymbDigWTotCorr")
```
```{r, include = FALSE}
analysisRaw %>% filter(is.na(SymbDigWTotCorr)) %>% select(PtID, SymbDigWTotCorr, SymbDigWNotDone)

analysisRaw <- analysisRaw %>% select(-SymbDigWNotDone)
```
PtIDs 24, 131, 150, 13, 7, 140, 149, 124: Symbol digit modalities test written was not done

### Symbol Digit Modalities Test - Oral

```{r}
getOverallStratifiedCounts(analysisRaw, "SymbDigOTotCorr")
```
```{r, include = FALSE}
analysisRaw %>% filter(is.na(SymbDigOTotCorr)) %>% select(PtID, SymbDigOTotCorr, SymbDigONotDone)

analysisRaw <- analysisRaw %>% select(-SymbDigONotDone)
```
PtIDs: 115, 24, 131, 150, 13, 7, 140, 149 113: Symbol Digit Modalities Test - Oral was not done

### Trail Making Test - Trail A

```{r}
getOverallStratifiedCounts(analysisRaw, "TrailMakATotTime")
```

This variable is complete

```{r, include = FALSE}
analysisRaw <- analysisRaw %>% select(-TrailMakANotDone)
```


### Train making Test - Trail B

```{r}
getOverallStratifiedCounts(analysisRaw, "TrailMakBTotTime")
```

```{r, include = FALSE}
analysisRaw %>% filter(is.na(TrailMakBTotTime)) %>% select(PtID, TrailMakBTotTime, TrailMakBNotDone)

analysisRaw <- analysisRaw %>% select(-TrailMakBNotDone)
```

PtIDs 56, 47, 144, 4, 159, 150: Trail Make B was not done

### Grooved Pegboard Test (dominant hand)

```{r}
getOverallStratifiedCounts(analysisRaw, "GrPegDomTotTime")
```
```{r, include = FALSE}
analysisRaw %>% filter(is.na(GrPegDomTotTime)) %>% select(PtID, GrPegDomTotTime, GrPegDomNotDone)

analysisRaw <- analysisRaw %>% select(-GrPegDomNotDone)
```

PtIDs 4, 132: Test was not done

### Geriatric depresssion scale

```{r}
analysisRaw %>%
  ggplot(aes(x = geriDepressionScore)) +
  geom_histogram(bins = 16)
```

```{r, include = FALSE}
analysisRaw %>% filter(is.na(geriDepressionScore)) %>%
  select(PtID, geriDepressionScore)
```
Incomplete score for PtID 80 (missing scores for 4 items)


### Diabetes numeracy test

Need to know how to score this
```{r}
analysisRaw %>%
  ggplot(aes(x = diabNumeracyScore)) +
  geom_histogram(bins = 16)
```

```{r, include = FALSE}
analysisRaw %>% filter(is.na(diabNumeracyScore)) %>% select(PtID, diabNumeracyScore)
```
Missing some number of items and thus a score for PtIDs: 4, 79, 175, 164, 159, 180, 153, 140, 60, 30, 32, 103,  121, 89, 83

### Functional activities questioinnaire

```{r}
getOverallStratifiedCounts(analysisRaw, "FuncActTotTestScore")
```
This variable is complete

```{r}
analysisRaw <- analysisRaw %>% select(-FuncActNotDone)
```


### Hypoglycemia fear survey

Need to know how to score this


### Duke social support

```{r}
getOverallStratifiedCounts(analysisRaw, "DukeSocTotDSSI")
```

```{r, include = FALSE}
analysisRaw %>% filter(is.na(DukeSocTotDSSI)) %>% select(PtID, DukeSocNotDone, DukeSocTotDSSI)

analysisRaw <- analysisRaw %>% select(-DukeSocNotDone)
```

PtID 124: Duke social score was not done

### Frailty

Taking the average of the measured frailty walks

```{r}
analysisRaw <- analysisRaw %>% 
  mutate(FrailtyFirstWalkTotTimeMin = FrailtyFirstWalkTotTimeMin*60,
         FrailtySecWalkTotTimeMin = FrailtySecWalkTotTimeMin*60,
         frailty1 = FrailtyFirstWalkTotTimeMin + FrailtyFirstWalkTotTimeSec,
         frailty2 = FrailtySecWalkTotTimeMin + FrailtySecWalkTotTimeSec) %>%
  rowwise() %>%
  mutate(frailty = mean(c(frailty1, frailty2), na.rm = TRUE)) %>%
  select(-c(FrailtyFirstWalkTotTimeMin, FrailtyFirstWalkTotTimeSec, 
            FrailtySecWalkTotTimeMin, FrailtySecWalkTotTimeSec, frailty1, frailty2))

analysisRaw %>%
  ggplot(aes(x = frailty)) +
  geom_histogram(binwidth = 0.5)
```
```{r, include = FALSE}
analysisRaw %>% filter(is.na(frailty))
```
PtIDs 14 and 112 missing frailty score

### Hypoglycemia unawareness

```{r}
getOverallStratifiedCounts(analysisRaw, "hypoUnaware")
```
```{r, include = FALSE}
analysisRaw %>% filter(is.na(hypoUnaware))
```
PtIDs 144, 158, 80 are missing hypoglycemia awareness


```{r}
names(analysisRaw)
analysisRaw <- analysisRaw %>% select(-c(FreqHypoDamage, DangersHighBG, VisitDaysFromEnroll, BGVisit1NotDone,
                          T1DDiagAge, InsUnkown, DealHypoEp, UndertreatHypo, HighBGDamage,
                          nInsurance, NumMeterCheckDayUnk, BBGAttitudeScaleNotDone, BGVisit1NotDone)) 
analysisObs <- analysisRaw %>% filter(str_detect(exclude, "No")) 

analysisObs[complete.cases(analysisObs), ] %>%
  group_by(BCaseControlStatus) %>%
  count()

analysisRaw %>% write_csv("../2_pipeline/2_analysisWithMissing.csv")
```

